---
title: "Project 2"
author: "Cassandra Chou"
date: "2024-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1
### part a)
```{r part 1a}
library(tidyverse)
library(tidytuesdayR)

Exp <- function(x, k) {
  exp_val <- 1
  for (i in 1:k) {
    exp_val <- exp_val + x^i/factorial(i)
  }
  exp_val
}

Exp(3, 2)
Exp(3, 100)

```
### part b)
```{r part 1b}
sample_mean <- function(x) {
  n <- length(x)
  mean_val <- 0
  for (i in 1:n) {
    mean_val <- mean_val + x[i]
  }
  
  mean_val/n
}

sample_sd <- function(x) {
  n <- length(x)
  sd_val <- 0
  mean_x <- sample_mean(x)
  
  for (i in 1:n) {
    sd_val <- sd_val + (x[i] - mean_x)^2
  }
  
  sd_val <- 1/(n-1) * sd_val
  sqrt(sd_val)
}

sample_mean(c(1, 2, 3, 4, 5))
sample_sd(c(1, 2, 3, 4, 5))

sample_mean(c(1, 4, 6, 5, 8, 1))
sample_sd(c(1, 4, 6, 5, 8, 1))

```


### part c)
```{r part1c}
calculate_CI <- function(x, conf = 0.95) {
  alpha <- 1 - conf
  degrees_freedom <- length(x) - 1
  t_score <- qt(p = alpha / 2, df = degrees_freedom, lower.tail = FALSE)
  
  mean_x <- mean(x)
  se_x <- sd(x)/sqrt(length(x))
  lower_bound <- mean(x) - t_score*se_x
  upper_bound <- mean(x) + t_score*se_x
  
  c(lower_bound, upper_bound)
}

# Test
x <- c(1:100)
dat <- data.frame(x = x)
fit <- lm(x ~ 1, dat)

# Calculate a 95% confidence interval
confint(fit, level = 0.95)

calculate_CI(x)

# Calculate a 90% confidence interval
confint(fit, level = 0.9)
calculate_CI(x, 0.9)

```

## Part 2
```{r part2 setup}
library(here)
if (!file.exists(here("data", "tuesdata_rainfall.RDS"))) {
    tuesdata <- tidytuesdayR::tt_load("2020-01-07")
    rainfall <- tuesdata$rainfall
    temperature <- tuesdata$temperature

    # save the files to RDS objects
    saveRDS(tuesdata$rainfall, file = here("data", "tuesdata_rainfall.RDS"))
    saveRDS(tuesdata$temperature, file = here("data", "tuesdata_temperature.RDS"))
}

rainfall <- readRDS(here("data", "tuesdata_rainfall.RDS"))
temperature <- readRDS(here("data", "tuesdata_temperature.RDS"))

library(tidyverse)
library(dplyr)

glimpse(rainfall)
glimpse(temperature)

```

```{r part2}
df <- rainfall %>% drop_na() %>% mutate(date = ymd(paste(year, month, day, sep = "-"))) %>% select(-month, -day)
df$city_name <- str_to_upper(df$city_name)

df <- inner_join(df, temperature, by = c("city_name",  "date"), relationship =
  "many-to-many")

glimpse(df)

```

## Part 3
### part a)
```{r part 3a}
library(ggplot2)
library(extrafont)

ggplot(data = df %>% filter(year >= 2014), aes(x = date, y = temperature)) + geom_point(aes(color = temp_type)) + facet_wrap(~city_name, scales = "free_y") + labs(title = "Temperature over Time, Sorted by City", subtitle = "Maximum and Minimum Temperatures Fluctate During the Year Across All Cities, \nPeaking Around the Beginning of the Year", caption = "Data Sourced from Tidy Tuesday", color = "Temperature Type") + xlab("Year") + ylab("Temperature (in Celsius)") + theme(plot.title = element_text(family = "Segoe UI Black", size = 16), axis.title.x = element_text(family = "Segoe UI Semilight"), axis.title.y = element_text(family = "Segoe UI Semilight"), plot.subtitle = element_text(family = "Segoe UI Semilight"), plot.caption = element_text(family = "Segoe UI Semilight", size = 11), panel.background = element_rect(fill = "#fff1da"))
  
  

```


### part b)
```{r part 3b}
plot_rainfall <- function(city_name, year) {
  city_nm <- str_to_upper(city_name)
  yr <- year
  
  if (city_nm %in% df$city_name == FALSE) {
    stop("Invalid City!")
  }
  
  if (yr %in% df$year == FALSE) {
    stop("Invalid Year!")
  }
  
  df_city_year <- df %>% filter(city_name == city_nm, year == yr)
  
  if(count(df_city_year) == 0) {
    stop("No combinations of this city/year!")
  }
  
  plot_title <- paste("Rainfall in", str_to_title(city_nm), "During the Year", yr, sep = " ")
  plot_subtitle <- "Exploring the Distribution of Rainfall on a Log Scale"
  
  ggplot(df_city_year, aes(log(rainfall))) +
  geom_histogram(binwidth = 0.3, fill = "#094b7c") + xlab("Log(Rainfall)") + ylab("Count of Rainfall Events") + labs(title = plot_title, subtitle = plot_subtitle, caption = "Data Sourced from Tidy Tuesday") +
  theme(plot.title = element_text(family = "Segoe UI Black", size = 16), plot.subtitle = element_text(family = "Segoe UI Semilight"), axis.title.x = element_text(family = "Segoe UI Semilight"), plot.caption = element_text(family = "Segoe UI Semilight", size = 11), axis.title.y = element_text(family = "Segoe UI Semilight"), panel.background = element_rect(fill = "#c3edff"))
}

plot_rainfall("PERTH", 2000)
plot_rainfall("MELBOURNE", 1999)

```

Function Design: First, I convert the city name to all captials so that if someone typed "Perth" or "perth" it would still be a valid city name. Then, I check to see if the city and year are valid. If not, I throw an error. Then, I filter the data, and check to see that there is at least one point in this city/year combination. If not, I throw an error saying that this city/year combination is not valid. Then, I create the plot title, and create the plot. I relabeled the axes and the title for clarity and interpretability. I added a bin width in case of city/year combinations with varying dataset sizes, and some color to the plot for aesthetics.

## Part 4
### part a)
```{r part 4a}
rain_df <- df %>% filter(year >= 2014) %>% group_by(city_name, year) %>% drop_na() %>% summarise(mean = sample_mean(rainfall), sd = sample_sd(rainfall), lower_bound = calculate_CI(rainfall)[1], upper_bound = calculate_CI(rainfall)[2])

rain_df

```

```{r part 4b}
rain_df %>% ggplot(aes(x = year, y = mean)) + geom_point() + geom_line() + geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.3) + facet_wrap(~city_name) + xlab("Year") + ylab("Mean Amount of Rainfall (mm)") + labs(title = "Mean Amounts of Rainfall in Australia from 2014-2019", subtitle = "Trends in Five Austrailian Metropolitan Cities with Fixed Scale Axes\nHigher Amounts of Rain in Melbourne and Perth, with Overall Decreasing Amounts of Mean Rainfall", caption = "Data Sourced from Tidy Tuesday") + theme(plot.title = element_text(family = "Segoe UI Black", size = 16), axis.title.x = element_text(family = "Segoe UI Semilight"), plot.caption = element_text(family = "Segoe UI Semilight", size = 11), axis.title.y = element_text(family = "Segoe UI Semilight"), plot.subtitle = element_text(family = "Segoe UI Semilight"), panel.background = element_rect(fill = "#e5f5ff"), strip.background = element_rect(fill = "#c9d3d9"))

rain_df %>% ggplot(aes(x = year, y = mean)) + geom_point() + geom_line() + geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.3) + facet_wrap(~city_name, scales = "free_y") + xlab("Year") + ylab("Mean Amount of Rainfall (mm)") + labs(title = "Mean Amounts of Rainfall in Australia from 2014-2019", subtitle = "Trends in Five Austrailian Metropolitan Cities with City-Defined Axes\nMostly fluctuating amounts of mean rainfall, with a slight trend towards decreasing amounts of rain", caption = "Data Sourced from Tidy Tuesday") + theme(plot.title = element_text(family = "Segoe UI Black", size = 16), axis.title.x = element_text(family = "Segoe UI Semilight"), plot.subtitle = element_text(family = "Segoe UI Semilight"), plot.caption = element_text(family = "Segoe UI Semilight", size = 11), axis.title.y = element_text(family = "Segoe UI Semilight"), panel.background = element_rect(fill = "#e5f5ff"), strip.background = element_rect(fill = "#c9d3d9"))

```

I plotted two plots with different y-scaling because it is important to see individualized trends in the data by city (city-defined axes), but it is also important to see how rainfall compares between the cities (fixed-scale axes).


